### 2022-11-28 ms.rmd
Doing some re-organisation and checking imputed values to real ones for cross valication

# functions
source("./R/func.R")
t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
data <- data %>% mutate(species = gsub(" ", "", species),
                        phylo = species,
                        obs = 1:n()) %>% filter(!species %in% t$Species_InData_But_NotTree) 
tree3 <- tree_checks(tree2, data = data, dataCol = "species", type = "prune")
tree_checks(tree3, data = data, dataCol = "species", type = "check")
# Grab phylo matrix
phy_vcv <- vcv(tree3, cor = FALSE)
tree3$node.label <- NULL

@@

```{r}

#### Try MCMCglmm which I expect will be MUCH faster. Deals with missing data using data augmentation
rerun = FALSE
if(rerun){
  prior = list(R = list(V = diag(3), nu=1.002), 
             G = list(G1=list(V = diag(3), nu = 1.002)))

@@

# We'll also z-transofmr to make the coprrelation stuff easier to calculate. 
data <- data %>% mutate(z_ctmax = scale(ctmax),
                         z_cfts = scale(cfts),
                        z_tdt_z = scale(tdt_z))

@@

# Extract estimates
sol <- mcmcglmm_model$Sol
vcv <- mcmcglmm_model$VCV

# Extract the estmated / imputaed Z for each species, but drop internal ancestral nodes
z <- sol[, grep("traittdt_z.animal", colnames(sol))]
z <- z[,-grep("Node", colnames(z))]
head(z)

# Clean up names so they are just species names

>>>>

# functions
  source("./R/func.R")
  t <- tree_checks(tree2, data = data, dataCol = "species", type = "check")
  data <- data %>% mutate(species = gsub(" ", "", species),
                          phylo = species,
                          obs = 1:n()) %>% filter(!species %in% t$Species_InData_But_NotTree) 
  tree3 <- tree_checks(tree2, data = data, dataCol = "species", type = "prune")
  tree_checks(tree3, data = data, dataCol = "species", type = "check")

  # Grab phylo matrix
  phy_vcv <- vcv(tree3, cor = FALSE)
  tree3$node.label <- NULL

@@

```{r, models}

### MCMCglmm Bayesian models. Deals with missing data using data augmentation

rerun = FALSE
if(rerun){
    prior = list(R = list(V = diag(3), nu=1.002), 
               G = list(G1=list(V = diag(3), nu = 1.002)))

@@

  # We'll also z-transform to make the correlation easier to calculate. 
  data <- data %>% mutate(z_ctmax = scale(ctmax),
                           z_cfts = scale(cfts),
                          z_tdt_z = scale(tdt_z))

@@

```{r, posteriors}

# Extract posterior estimates
  sol <- mcmcglmm_model$Sol
  vcv <- mcmcglmm_model$VCV

# Extract the estimated / imputed Z for each species, but drop internal ancestral nodes
  z <- sol[, grep("traittdt_z.animal", colnames(sol))]
  z <- z[,-grep("Node", colnames(z))]
  head(z)

# Clean up names so they are just species names
  colnames(z) <- gsub("traittdt_z.animal.", "", colnames(z))

# Have a look at how well the model gets the actual data correct.
  z_known <- data %>% filter(!is.na(tdt_z)) %>% dplyr::select(species, tdt_z) %>% distinct()
    z_est <- apply(z, 2, function(x) median(x))[which(names(z_est) %in% z_known$species)]
